A Combinatorial Search of Parameterized Quantum Circuit Learning for Chemical Applications

### 2025-01-20 manuscript/main.tex
manuscript Mon 20 Jan 2025 17:04:40 EST

\subsection{Dataset}
In this study, we explore three datasets: a function fitting dataset, consisting of a noisy linear, quadratic, and sine function, used for model calibration; a dataset consisting of electronic structure features to predict wavefunctions using the data-driven coupled-cluster scheme of Townsend and Vogiatzis \cite{townsend_data-driven_2019}; and a dataset of bond separation energies (BSE) of molecules, where the feature set encodes structural information of each molecule.

The noisy function fitting dataset is generated by
$f(x) = x + \varepsilon$ such that $\varepsilon \in \mathcal{N} (0, \sigma^{2} )$, where $\sigma\approx0.34$. The linear function is $f(x)=x$, quadratic is $f(x) = 0.3967 x^{2} - 3.2807 x + 6.7569$, and sine function is $f(x) = 0.9913 \sin(2.3838 x+0.0344)$. 
The normalized versions are shown in Fig. \ref{fig:train_vs_test}.

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{../images/Function_Fitting/function_dataset/linear_train_vs_test.png}
		\caption{}
		\label{fig:linear_train_vs_test}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{../images/Function_Fitting/function_dataset/quadratic_train_vs_test.png}
		\caption{}
		\label{fig:quadratic_train_vs_test}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{../images/Function_Fitting/function_dataset/sine_train_vs_test.png}
		\caption{}
		\label{fig:sine_train_vs_test}
	\end{subfigure}
	\caption{}
	\label{fig:train_vs_test}
\end{figure}

\cite{rupp_fast_2012}
\cite{rogers_extended-connectivity_2010}

The first chemical dataset we explore is the BSE49 dataset which consists of homolytic bond cleavages, such as \ce{A-B -> A^{.} + B^{.}}, of X molecules with 49 unique bond types.
One important aspect of this dataset is the choice of molecular representation, or how the molecule is represented in the machine learning models.
To find the optimal molecular representation, we examined five different molecular representations including Coulomb Matrices \cite{rupp_fast_2012}, Molecular ACCess Systems fingerprint (MACCS)\cite{durant_reoptimization_2002}, persistence images (PIs) \cite{adams_persistence_2017,townsend_representation_2020,schiff_augmenting_2022}, RDKit [cite], and Morgan or extended-connectivity fingerprints \cite{morgan_generation_1965,rogers_extended-connectivity_2010}.
For each representation we tested two different choices of representation, one based on the the products minus the reactants, denoted by \textit{sub}, similar to the method used in [cite], and one based on just the reactants, denoted as \textit{AB}.
We use the same 80\% train/20\% test split as we use throughout our studies. 
We found that the best molecular representation across all models test, as shown in Fig. \ref{fig:classicalfuncfit}, was Morgan fingerprints using the \textit{sub} formulation.

>>>>

\subsection{Datasets}
In this study, we explore three datasets: a function fitting dataset, consisting of a noisy functions used for model calibration; a dataset consisting of electronic structure features to predict wavefunctions using the data-driven coupled-cluster scheme of Townsend and Vogiatzis \cite{townsend_data-driven_2019}; and a dataset of bond separation energies (BSE) of molecules, where the feature set encodes structural information of each molecule. 
The function fitting dataset consists of a noisy linear (Fig. \ref{fig:linear_train_vs_test}), quadratic (Fig. \ref{fig:quadratic_train_vs_test}), and sine (\ref{fig:sine_train_vs_test}) function.  
For each function, 90 points are generated, where 80 are used for training set and 10 for the test set.
This dataset is used for calibration of the 168 PQCs, along with providing additional insights into the effects of circuit depth, training set size, and error mitigation on model performance.

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{../images/Function_Fitting/function_dataset/linear_train_vs_test.png}
		\caption{}
		\label{fig:linear_train_vs_test}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{../images/Function_Fitting/function_dataset/quadratic_train_vs_test.png}
		\caption{}
		\label{fig:quadratic_train_vs_test}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.3\textwidth}
		\centering
		\includegraphics[width=\textwidth]{../images/Function_Fitting/function_dataset/sine_train_vs_test.png}
		\caption{}
		\label{fig:sine_train_vs_test}
	\end{subfigure}
	\caption{The function fitting dataset which consists of 90 total points, 80 for training and 10 used for testing, for the (a) linear, (b) quadratic, and (c) sine functions.}
	\label{fig:train_vs_test}
\end{figure}

Following model calibration using the function fitting dataset, we explore the appliciblity of PQCs for complex chemically relevant machine learning tasks.
The first chemically motivated dataset we explore is the BSE49 dataset, which contains the bond seperation energies (BSE)  for the homolytic bond cleavage of covalently bonded molecules, such as \ce{A-B -> A^{.} + B^{.}}.\cite{prasad_bse49_2021}
This dataset consists of 4394 datapoints (!VIKI! paper says 4502), 1951 of which are existing and 2443 are hypthoetical structures, with 49 unique A-B single bond types.
In practice, we used 2436 of the hypothetical structures due to issues with valency exceptions when converting to RDKit mol objects which were later used for generating our features for the machine learning models.
An important aspect of machine learning in chemistry is the choice of molecular representation, or how the molecule is represented in the machine learning models.\cite{jones_molecular_2023}

 Cite:
DScribe\cite{himanen_dscribe_2020} CMs and SOAPs + REMatchKernel

Ripser\cite{tralie_ripserpy_2018}
RDKit\cite{noauthor_rdkit_nodate}

SOAPs\cite{de_comparing_2016} were excluded due to the cost of computing the kernel

To find the optimal molecular representation, we examined five different molecular representations including Coulomb Matrices \cite{rupp_fast_2012}, Molecular ACCess Systems fingerprint (MACCS)\cite{durant_reoptimization_2002}, persistence images (PIs) \cite{adams_persistence_2017,townsend_representation_2020,schiff_augmenting_2022}, RDKit [cite], and Morgan or extended-connectivity fingerprints \cite{morgan_generation_1965,rogers_extended-connectivity_2010}.

\begin{equation}
	X_{\text{RDKit},i} = X_{\text{RDKit},i}^{\text{R}} - X_{\text{RDKit},i}^{\text{R}}
	\label{eq:descriptor_garcia}
\end{equation}

An additional challenge for choosing a molecular representions for the BSE49 dataset is how to represent the overall chemical system

 We found that the best molecular representation across all models test, as shown in Fig. \ref{fig:classicalfuncfit}, was Morgan fingerprints using the \textit{sub} formulation.\cite{garcia-andrade_barrier_2023}

For each representation we tested two different choices of representation, one based on the the products minus the reactants, denoted by \textit{sub}, similar to the method used in [cite], and one based on just the reactants, denoted as \textit{AB}.

We use the same 80\% train/20\% test split as we use throughout our studies.
